### 2023-04-27 chapter2/chapter2.tex
Added paragraph about motifs as features

% This form of constraint is    B. Though other analyses have identified Though IDRs short linear motifs (SLiMs) are enriched in IDRs and are sometimes strongly conserved at specific positions, these residues constitute a small fraction of disordered regions. However, because these motifs are short and some are highly permissive at specific positions, they have high evolutionarily plasticity. Thus, motifs
%Though these motifs are sometimes conserved by position and can be identified as short segments of conservation in otherwise highly diverged alignments of IDRs, their brief lengths make them evolutionarily plastic. Thus, they are easily

@@

% IV. Explicit phylogenetic models of trait evolution can reveal patterns of evolution with fewer assumptions
%    A. Previous analyses compared observed distributions of features in observed IDRs to the distributions generated by simulated counterparts
%    B. Differences were interpreted as evidence of selection on features
%        1. However, evidence of deviation from the null model rather than evidence of constraint necessarily
%        2. Null distribution generated by a model of evolution where the composition of sequences stochastically evolve towards a specific equilibrium distribution
%        3. Previous studies have identified IDRs with varying compositional preferences, so a single equilibrium distribution is unlikely to globally describe observed patterns of sequence evolution in IDRs
%    C. Thus in this work we explore use of various statistical approaches that directly measure rates of change and constraint of molecular features indepedent of specific model of sequence evolution
%        1. Place these evolutionary analyses in a statistical phylogenetic framework through the application of the Brownian motion and Ornstein-Uhlenbeck models of trait evolution.
%        2. Though these models have own assumptions, they are macroscopic descriptions of continuous trait evolution and therefore independent of any assumptions about evolution at the level sequences
%        3. Still have their own assumptions of course, but more general and robust to deviations
%        4. We view these approaches as complementary, each having different trade-offs between specificity and generality of testing specific hypotheses of disorder evolution
%        5. We also explore hybrid approaches where the evolutionary rates obtained from phylogenetic models are used as a statistic
%
% V. Therefore, in this work we apply evolutionary models of sequence and trait evolution to the sequences of IDRs and their of associated molecular feature to dissect the forces driving their evolution. We find IDRs exhibit unique patterns of evolution at the level of sequences. We find intrinsic disorder is a dynamically evolving property in some regions, and changes in disorder preference are correlated with changes in the complexity. We find that IDRs are unconstrained along several axis of trait evolution, but they also display

>>>>

This form of selective constraint can also describe the evolution of more ``localized'' features in IDRs such as short linear motifs (SLiMs). Because SLiMs are composed of fewer than 12 residues, they form limited interfaces that frequently mediate the transient binding events involved in signaling pathways~\cite{Tompa2014}. Accordingly, they are highly enriched in IDRs, which provide an accessible and flexible scaffold for these interactions~\cite{Fuxreiter2007, Davey2012}. While some SLiMs in IDRs are strongly conserved at specific positions, these constitute a small fraction of disordered residues, estimated at roughly 17\% in the yeast proteome~\cite{NguyenBa2012}. Instead, as SLiMs are compact and often highly degenerate at some positions, they can arise de novo from a small number of mutations and therefore have high rates of turnover. Furthermore, when IDRs contain multiple copies of a motif that jointly mediate a high-avidity interaction~\cite{Tompa2014} or a graded response to a signal via the accumulation of multiple phosphorylations~\cite{Wright2014}, the individual motifs are under weak selective constraints. As a result, though SLiMs are encoded by specific sequences, in some contexts they may evolve as distributed features that characterize IDRs as a whole rather than specific sites within them.

@@

% IV. Thus, in this work Explicit phylogenetic models of trait evolution can reveal patterns of evolution with fewer assumptions
%    A. Previous analyses compared observed distributions of features in observed IDRs to the distributions generated by simulated counterparts
%    B. Differences were interpreted as evidence of selection on features
%        1. However, evidence of deviation from the null model rather than evidence of constraint necessarily
%        2. Null distribution generated by a model of evolution where the composition of sequences stochastically evolve towards a specific equilibrium distribution
%        3. Previous studies have identified IDRs with varying compositional preferences, so a single equilibrium distribution is unlikely to globally describe observed patterns of sequence evolution in IDRs
%    C. Thus in this work we explore use of various statistical approaches that directly measure rates of change and constraint of molecular features indepedent of specific model of sequence evolution
%        1. Place these evolutionary analyses in a statistical phylogenetic framework through the application of the Brownian motion and Ornstein-Uhlenbeck models of trait evolution.
%        2. Though these models have own assumptions, they are macroscopic descriptions of continuous trait evolution and therefore independent of any assumptions about evolution at the level sequences
%        3. Still have their own assumptions of course, but more general and robust to deviations
%        4. We view these approaches as complementary, each having different trade-offs between specificity and generality of testing specific hypotheses of disorder evolution
%        5. We also explore hybrid approaches where the evolutionary rates obtained from phylogenetic models are used as a statistic
%
% V. Therefore, in this work we apply evolutionary models of sequence and trait evolution to the sequences of IDRs and their of associated molecular feature to dissect the forces driving their evolution. We find IDRs exhibit unique patterns of evolution at the level of sequences. We find intrinsic disorder is a dynamically evolving property in some regions, and changes in disorder preference are correlated with changes in the complexity. We find that IDRs are unconstrained along several axis of trait evolution, but they also display
